---
title: "Final Project Group 9"
subtitle: "SOC 225: Data & Society, Spring 2020"
author: "Jingzhao Huang, Lauren Liao, Wei Yang, Angela Zhao, Siyue Zhu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---


```{r}
library(tidyverse)
library(dplyr)
```


# 1: Data Collection

## Import datasets

```{r}
lifeExp_data <- read_csv("data/Life Expectancy Data.csv")
continent_data <- read_csv("data/Countries-Continents.csv")
```

## Provide summary statistics for related variables

```{r}
summary(lifeExp_data$`Life expectancy`)
summary(lifeExp_data$GDP)
```


# 2: Data Preparation


## Merge two sources of data + clean NAs

```{r}
whole_data <- left_join(lifeExp_data, continent_data, by = c("Country"="Country"), na = c("NA", "N/A"))
```

## Rename the colomn `GDP` as `GDP_per_capita`

```{r}
names(whole_data)[names(whole_data) == "GDP"] <- "GDP_per_capita"
```

## Clean other missing values

```{r}
whole_data <-
  whole_data %>% 
  filter(`Life expectancy` != 0 & GDP_per_capita != 0)
```

## Create a numeric variable: GDP

```{r}
whole_data <-
  whole_data %>% 
  mutate(GDP = GDP_per_capita * Population)
``` 

## Provide summary statistics for revised dataset

```{r}
summary(whole_data)
glimpse(whole_data)
```


# 3: Visual Analysis

## Histogram for Life Expectancy in All Countries

```{r}
ggplot(whole_data, aes(x = `Life expectancy`)) +
  geom_histogram() +
  labs(title = "Life Expectancy in All Countries")
```


## Histogram for GDP per capita in All Countries

```{r}
whole_data %>% 
  drop_na() %>%
  filter(GDP_per_capita != 0) %>%
  ggplot(aes(x = GDP_per_capita)) +
  geom_histogram() +
  labs(title = "GDP per capita in All Countries")
```


## Scatter Plot between Life Expectancy and GDP per capita in All Countries

```{r}
ggplot(whole_data, aes(x = GDP_per_capita, y = `Life expectancy`)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Life Expectancy and GDP per capita in All Countries")
```


## Boxplot for Life Expectancy in Developing and Developed Countries

```{r}
ggplot(whole_data, aes(x = Status, y = `Life expectancy`, color = Status)) + 
  geom_boxplot() + 
  labs(title = "Boxplot for Life Expectancy in Developing and Developed Countries")
```


## Scatter Plot between Life Expectancy and GDP in Developing Countries 

```{r}
developing_data <- filter(whole_data, Status == 'Developing')
ggplot(developing_data, aes(x = GDP_per_capita, y = `Life expectancy`)) + 
  geom_point() + 
  geom_smooth(method = "lm", colour = "red") + 
  guides(color = FALSE) + 
  labs(title = "GDP per capita and Life Expectancy in Developing Countries")
```


## Scatter Plot between Life Expectancy and GDP per capita in Developed Countries

```{r}
developed_data <- filter(whole_data, Status == 'Developed')
ggplot(developed_data, aes(x = GDP_per_capita, y = `Life expectancy`)) + 
  geom_point() + 
  geom_smooth(method = "lm", colour = "red") + 
  guides(color = FALSE) +
  labs(title = "GDP per capita and Life Ecpectancy in Develped Countries")
```

